---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

unzip file, read file with read.csv, check file structure, and change variable date to Date format

```{r}
unzip(zipfile = "activity.zip")
activity <- read.csv("activity.csv")
str(activity)
activity$date <- as.Date(activity$date, "%Y-%m-%d")

```


## What is mean total number of steps taken per day?
calculate daily steps using tapply, plot with histogram, get the mean and median value

```{r, fig.width= 5}
daily_step <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
hist(daily_step, breaks = 10, col = "skyblue1")
noquote("Mean of total steps per day is: ") 
mean(daily_step)
noquote("Median of total steps per day is: ") 
median(daily_step)

```

## What is the average daily activity pattern?
calculate averaged steps per interval with tapply, then averaged by days

make a data frame with interval and averaged steps

plot with line and find the max steps with which.max
```{r}
daily_pattern <- tapply(activity$steps, activity$interval, sum, na.rm = TRUE)
daily_pattern <- data.frame(names(daily_pattern), daily_pattern/length(daily_step))
names(daily_pattern) <- c("Time", "Steps")
plot(daily_pattern$Time, daily_pattern$Steps, type = "l", xlab = "Time", ylab = "Steps", xaxt = 'n', col = "skyblue2")
axis(1, at = seq(0, 2400, by = 400), labels = c("0:00", "4:00", "8:00", "12:00", "16:00", "20:00", "24:00"))
noquote("The interval contains the max averaged number of steps:")
daily_pattern[which.max(daily_pattern$Steps), 1]


```
## Imputing missing values
First find which columns contain NA value

Replace NA vaule with mean value

plot with histogram and recalculate mean and median
```{r, fig.width = 5}
library(dplyr)
list_na <- colnames(activity[apply(activity, 2, anyNA)])
list_na
noquote("The total number of missing value in the dataset:")
sum(is.na(activity[, colnames(activity) %in% list_na]))
mean_missing <- mean(activity$steps, na.rm = TRUE)
mean_missing
new_activity <- activity %>% mutate(steps = ifelse(is.na(steps), mean_missing, steps))
daily_step1 <- tapply(new_activity$steps, new_activity$date, sum, na.rm = TRUE)
hist(daily_step1, breaks = 10, col = "deepskyblue1", xlab = "daily steps")
noquote("Mean of total steps per day is: ") 
mean(daily_step1)
noquote("Median of total steps per day is: ") 
median(daily_step1)
imput <- ifelse(mean(daily_step1) > mean(daily_step), "increases", "decreases")
noquote(paste(c("Imputing missing values", imput, "the total daily number of steps."), sep = " ", collapse = " "))

```


## Are there differences in activity patterns between weekdays and weekends?
load lattice library

mutate weekdays with weekdays(date)

calculate average steps based on interval

plot with xyplot, by weekdays, using layout to get 2 vertical panels, using type "l" to plot line
```{r, fig.width = 5}
library(lattice)
new_activity <- new_activity %>% mutate(weekdays = ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
new_activity1 <- new_activity %>% group_by(interval, weekdays) %>% summarise(average_steps = mean(steps))
xyplot(average_steps ~ interval | weekdays, new_activity1, layout = c(1, 2), type = "l", ylab = "Number of steps")

```